---
title: 'Background'
output: md_document
---
```{r, echo = FALSE, message = FALSE}
library(data.table)
library(ggplot2)

devtools::load_all('..')

theme_set(
  theme_bw() +
    theme(panel.border = element_rect(color = NA))
)
knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE,
  fig.path = 'figure/', fig.align = 'center'
)
```

## Background

A _bank holding company_ (BHC) is simply a firm that controls one or more banks. I've made an attempt here to visualize the structures of such firms, with an eye toward portraying the scale and complexity of organization. The underlying data comes from the <a href='https://www.ffiec.gov/nicpubweb/nicweb/nichome.aspx' target='_blank'>National Information Center</a> (NIC) website, where holding company data can be queried via the Institution Search form. While the NIC reports complete data for U.S. companies only, some of them control a large number of international subsidiaries based in Europe, Asia, and elsewhere. Some holding companies are themselves U.S. subsidiaries of foreign banking and financial organizations such as UBS, Barclays, and Credit Suisse. Together, these companies control over $15 trillion in assets.

<hr>

**Note:** The data includes not only BHCs but also Financial Holding Companies, Savings & Loan Holding Companies, Intermediate Holding Companies, and certain Foreign Banking Organizations which have U.S. operations and are categorized as BHCs. These are all monitored by the Federal Reserve and are required to provide the structural data discussed below. See <a href='https://www.ffiec.gov/nicpubweb/content/help/institution%20type%20description.htm' target='_blank'>here</a> for definitions of these types.

<hr>

The figure below shows the size (measured by total assets) of the 20 largest HCs as of the end of 2016, along with their share of assets among the ~600 HCs with at least $1 billion in assets:

```{r, fig.width = 8 * 8 / 7, fig.height = 8, out.width = '800px', out.height = '700px', dpi = 150}
hc10bn <- fread('../data/HC10bn.csv')
setnames(hc10bn, 'Institution Name (RSSD ID)', 'name')
setnames(hc10bn, '12/31/2016 Total Assets', 'asset')
# Note: Assets in thousands of dollars
hc10bn[, asset:= as.numeric(gsub('[^0-9]', '', asset)) / 1e6]
hc10bn[, name:= gsub(' \\(\\d+\\)', '', name)]
hc10bn[, name:= gsub('ASSOCIATION', 'ASSOC.', name)]
hc10bn[, name:= factor(name, name)]
hc10bn[, cumAsset:= cumsum(asset) / TOTAL_FRY9C_ASSETS_BN_123116]

p <- ggplot(hc10bn[1:20], aes(y = asset, x = name)) +
  geom_bar(stat = 'identity', fill = 'royalblue') +
  labs(x = '', y = '12/31/2016 Total Assets (billions)') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.margin = unit(c(.6, 1.4, 0, 1.4), 'cm'))

p + geom_line(aes(x = 1:20, y = cumAsset * get_ymax(p)), lty = 2, lwd = 1.3, col = 'red') +
  scale_y_continuous(sec.axis = sec_axis(~ . / get_ymax(p), 'Cum. Share of Assets',
                                         breaks = seq(0, 1, .25)))

```

There are a large number of medium- and small-sized holding companies that control less than $1 billion in assets. The figure below shows the total number of registered holding companies by type, with annotations of significant legislation. (More about this...). The data comes from scraping <a href='https://www.ffiec.gov/nicpubweb/nicweb/InstitutionHistory.aspx?parID_RSSD=1021628&parDT_END=99991231' target='_blank'>Institution Histories</a> on the NIC website, which can be used to identify when a firm was organized as a holding company. (A master list of firms, domestic and foreign, that were ever holding companies is available <a href='https://www.chicagofed.org/banking/financial-institution-reports/bhc-name-list' target='_blank'>here</a>.)

```{r, fig.width = 8 * 8 / 6, fig.height = 8, out.width = '800px', out.height = '600px', dpi = 150}
HCcounts <- fread('../data/HCcounts.csv')
HCcounts[, yearqtr:= as.Date(yearqtr)]
HCcounts[, Type:= factor(Type, levels = names(BHC_CATEGORIES))]

ggplot(HCcounts, aes(x = yearqtr, y = V1)) +
  geom_area(aes(fill = Type), color = 'lightgray', position = 'stack',
            size = .2, alpha = .9) +
  scale_x_date(breaks = seq.Date(RMD_PLOT_START_DATE, RMD_PLOT_END_DATE, by = '10 years'),
               labels = seq(1960, 2010, 10)) +
  geom_vline(xintercept = as.numeric(BHC_AMEND_DATE), lty = 2) +
  annotate('text', x = BHC_AMEND_DATE + 240, y = 6000,
           label = 'Dec 1970:\nAmendments to BHCA',
           family = 'mono', size = 3, hjust = 0, vjust = 1) +
  geom_vline(xintercept = as.numeric(GLB_ACT_DATE), lty = 2) +
  annotate('text', x = GLB_ACT_DATE + 240, y = 6000,
           label = 'Nov 1999:\nGramm-Leach-Bliley Act',
           family = 'mono', size = 3, hjust = 0, vjust = 1) +
  labs(x = '', y = 'Number of holding companies')
```

The plot below decomposes the quarterly churn into different event types described in the Institution Histories: establishments, acquisitions, closures, and changes in organization type. I've removed the large "jump" following the passage of the 1970 BHCA amendments, when over 900 companies registered as BHCs:

```{r, fig.width = 8 * 8 / 6, fig.height = 8, out.width = '800px', out.height = '600px', dpi = 150}
HCevents <- fread('../data/HCevents.csv')
HCevents[, yearqtr:= as.Date(yearqtr)]
HCevents[, `Event Type`:= factor(`Event Type`, levels = BHC_EVENT_LEVELS)]

ggplot(HCevents, aes(x = yearqtr, y = value)) +
  stat_smooth(geom = 'area', aes(fill = `Event Type`), position = 'stack',
              color = 'lightgray', span = .06, size = .2, alpha = .9) +
  scale_fill_manual(values = CHURN_PLOT_COLORS) +
  # Total (net) churn
  geom_smooth(data = HCevents[, .(value = sum(value)), by = 'yearqtr'],
              aes(x = yearqtr, y = value, color = 'Total churn'), lty = 2, span = .06,
              lwd = .5, se = FALSE) +
  scale_color_manual('', values = 'black') +
  scale_x_date(breaks = seq.Date(RMD_PLOT_START_DATE, RMD_PLOT_END_DATE, by = '10 years'),
               labels = seq(1960, 2010, 10)) +
  scale_y_continuous(limits = c(NA, 250)) +
  labs(x = '', y = 'Number of events')
```

Now on to the structure data, which is the primary focus of this application;

<hr>

**Note:** The structure data is not easy to access. For large institutions (i.e., the most interesting ones) the NIC search form returns only pdf documents, which must be processed to obtain usable data structures. Some regional Federal Reserve banks publish <a href='https://www.richmondfed.org/banking/supervision_and_regulation/fry6_reports' target='_blank'>FR Y-6 filings</a> with more detailed information about holding company structures, but as "unusable" pdf scans.

<hr>

The company structures are not strict hierarchies, in the sense that subsidiaries may be partially controlled by multiple parent companies (this is not uncommon). The definition of "control" (say, of company B by company A) is established in the instructions for the <a href='https://www.federalreserve.gov/apps/reportforms/reportdetail.aspx?sOoYJ+5BzDaGhRRQo6EFJQ==' target='_blank'>FR Y-10 Reporting Form</a> and includes such criteria as

* Ownership of 25% or more of any class of B's voting securities;

* Election of a majority of B's board of directors, trustees, general partners, or other high-level management positions;

* Ownership of "all or substantially all" of B's assets;

* ...
	
The forms also provide information on entity types and locations. Most entities are given nondescript labels of either "Domestic Entity Other" or "International Nonbank Subsidiary of Domestic Entities"; others have more precise labels such as "Data Processing Servicer", "Securities Broker/Dealer", "Federal Savings Bank" and "Edge Corporation." (A glossary describing each of these types is available <a href='https://www.ffiec.gov/nicpubweb/content/help/institution%20type%20description.htm' target='_blank'>here</a>.) To simplify visualization, I've aggregated the types into eight categories as shown in the network legend. (A "nonbank", while classified as such, may engage in banking-related activities.) An entity's "location" refers to its physical location as reported in the FR Y-10.

<hr>

**Note:** While holding companies report detailed industry classifications (NAICS) of each of their subsidiaries in the reporting forms, this information is not publicly available through the NIC website. (Although, in many cases the general activity is sort of evident from the entity's name.) The New York Fed research articles referenced have special access to this information and discuss it in some detail.

<hr>

### Research Questions
* How can this data be used to characterize the complexity of financial instutitions? (Is there a better measure than simply counting the number of controlled entities?)

* Besides simply being larger, how are the largest HCs structurally different from medium- and small-sized ones?

* What are the (regulatory) incentives for structuring these institutions a certain way?

* How has the structure of these institutions changed over time? (More complex?)

* What activities do non-banks engage in? (If this info is not available directly, as noted above, then maybe can discern from the names of entities.)

* How are offshore financial centers (OFCs) incorporated into the structures?

* ...


### References
Avraham, D., P. Selvaggi and J. Vickery. "A Structural View of U.S. Bank Holding Companies." FRBNY Economic Policy Review, July 2012. (<a href='https://www.newyorkfed.org/medialibrary/media/research/epr/12v18n2/1207avra.pdf' target='_blank'>link</a>)

Cetorelli, N., J. McAndrews and J. Traina. "Evolution in Bank Complexity." FRBNY Economic Policy Review, December 2014. (<a href='https://www.newyorkfed.org/medialibrary/media/research/epr/2014/1412cet2.pdf' target='_blank'>link</a>)

Cetorelli, N. and S. Stern. "Same Name, New Businesses: Evolution in the Bank Holding Company." FRBNY Liberty Street Economics blog, 28 September 2015. (<a href='http://libertystreeteconomics.newyorkfed.org/2015/09/same-name-new-businesses-evolution-in-the-bank-holding-company.html' target='_blank'>link</a>)

Flood, M., D. Kenett, R. Lumsdaine and J. Simon. "The Complexity of Bank Holding Companies: A Topological Approach." NBER Working Paper 23755, August 2017. (<a href=https://www.nber.org/papers/w23755 target='_blank'>link</a>)

<hr>

